---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ZikaModel

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/mrc-ide/ZikaModel.svg?branch=master)](https://travis-ci.org/mrc-ide/ZikaModel.svg?branch=master)
<!-- badges: end -->

`ZikaModel` is an R package for running the Zika transmission model 
developed at Imperial College London using R odin.

The transmission model is a metapopulation model which includes the 
dynamics of the _Aedes_ mosquito vector and the age-stratified human host 
populations. The model which has a stochastic and a deterministic version,
simulates also the impact of seasonality and control strategies, 
such as the release of Wolbachia-infected mosquitoes and child vaccinationon, 
on virus dynamics .

For details of the original transmission model please see the 
[Ferguson et al. 2016 paper](https://science.sciencemag.org/content/353/6297/353) 
which is the article where the model is published.

## Installation
You need to first install the [odin](https://github.com/mrc-ide/odin) R package. 

Once odin is installed, you can install the `ZikaModel` package with the 
following steps:  

- First install `devtools`, if you don't already have it

```{r include = TRUE, eval = FALSE}
install.packages("devtools")
library(devtools)
```
- Then, in a fresh R session, install the `ZikaModel` package

```{r include = TRUE, eval = FALSE}
devtools::install_github("mrc-ide/ZikaModel")
```

- Load and attach it 
```{r}
library(ZikaModel)
```

## Running the base model
Check this vignette on how to run the [deterministic](https://mrc-ide.github.io/ZikaModel/articles/deterministic_base_model.html) 
base version of the model. 
