---
title: "Run a customized model"
author: "Lorenzo Cattarino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run a custom model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ZikaModel)
```

## Customized models

`ZikaModel` allows the user to change the value of all the model parameters defined by `model_param_list_create()`. This allows more flexibility in the range of scenarios considered (e.g. inteventions to control the disease). Also, by taking advantage of this feature it is also possible to simulate the transmission dynamics of other immunizing single-strain infections transmitted by _Aedes aegypti_ mosquitoes, such as chikungunya. 

Here we will demonstrate how run a customized model which simulates vector control. different diseases  and diseasesThis requires to use a different set of functions which will be demonstrated here. There are three parameters which simulate this type of intevention:

* propMwtControl - Increase in mortality of adult wild type mosquitoes induced by the control
* TimeMwtControlOn - the time when the control starts
* TimeMwtControlOff - the time when the control ends

Let's assume that control increases mortality of adults mosquitoes by 20%. For now we will use the default values of `TimeMwtControlOn` and `TimeMwtControlOff`, which simulates control for one year starting on the second year of the simulation. To keep it simple we will assume no effect of seasonality across patches. This is analogous to run the model with only one patch.

```{r}
age_init <- c(1, 9, 10, 10, 10, 10, 10, 10, 10, 10, 10)

deathrt <- c(1e-10,
             1e-10,
             1e-10,
             0.00277068683332695,
             0.0210680857689784,
             0.026724997685722,
             0.0525354529367476,
             0.0668013582441452,
             0.119271483740379,
             0.279105747097929,
             0.390197266957464)

# number of years to run the simulation for
time_years <- 50 # years

# time step
my_dt <- 1

# number of model time steps
time_frame <- (364 * time_years) / my_dt

odin_model_path <- system.file("extdata/odin_model_determ.R", package = "ZikaModel")

propMwtControl <- 0.2

# this returns a function that will generate an instance of the model
create_generator <- create_r_model(odin_model_path = odin_model_path,
                                   agec = age_init,
                                   death = deathrt,
                                   nn_links = nn_links,
                                   amplitudes_phases = amplitudes_phases,
                                   DT = my_dt,
                                   season = FALSE,
                                   propMwtControl = propMwtControl)

# this generates an instance of the model using an initial equilibrium initialisation state 
gen <- create_generator$generator(user = create_generator$state)

integer_time_steps <- (364 * time_years) / my_dt

its <- seq(0, integer_time_steps, 1)

# run the model
mod_run <- gen$run(its)

# post processing
out <- gen$transform_variables(mod_run)
```

Now processing and plotting, as done before: 

```{r, fig.height = 4, fig.width = 6}
diagnostics <- post_processing(model_outputs, my_dt)

plot_compartments(diagnostics$compartments)
```

```{r, fig.height = 6, fig.width = 8}
plot_demographics(diagnostics$demographics)

mosquito_diagnostics <- post_processing_mos(model_outputs)

plot_demographics(mosquito_diagnostics)
```

```{r, fig.height = 6, fig.width = 8}
season_vars <- plot_Kc_eip_delta(model_outputs)
grid::grid.draw(season_vars)
```
