---
title: "Run a base model"
author: "Lorenzo Cattarino"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run a base model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Running the base model
Here we consider a deterministic model without seasonality and interventions. By defaut the model uses Brazil demographics (annual per capita human birth and death rates). This will likeley be modifed in future versions. We run the deterministic version of the model with `run_deterministic_model()`:

```{r}
library(ZikaModel)

r1 <- run_deterministic_model(time_period = 365 * 5)
```

The returned object is a `Zika_model_simulation` object, which is a list of two objects:

* `output` - model output
* `parameters` - model parameters

`Zika_model_simulation` objects can be plotted as follows:

```{r, fig.height = 6, fig.width = 8}
plot(r1, type = "H")
```

This will plot all the human compartments of the model, namely:

* `S` - Susceptibles 
* `I1` - Exposed and Infectious
* `R1` - Recovered

The argument `type` is used to specify which part of the model we want to plot, the human (`H`) or the mosquito (`M`) part. We can select which compartment to plot with the `var_select` argument. Arguments passed to `var_select` must be one of the variables in the above plot. 

```{r, fig.height = 6, fig.width = 8}
plot(r1, type = "H", var_select = c("S", "I1"))
```

Alternatively one of the following variables can be selected: 

* `Ntotal` - total size of the human population  
* `inf_1` - daily incidence of infections
* `MC` - daily incidence of microcepahly cases 
* `inf_1_w` - weekly incidence of infections per 1000 individuals 
* `MC_w` - weekly incidence of microcepahly cases per 1000 individuals.

First, we plot the weekly number of infections:

```{r, fig.height = 6, fig.width = 8}
plot(r1, type = "H", var_select = "inf_1_w")
```

and then the weekly number of microcephaly cases: 

```{r, fig.height = 6, fig.width = 8}
plot(r1, type = "H", var_select = "MC_w")
```

By default the `plot` function uses internally the `format_output_H()` or the `format_output_M()` function, depending on the value of the `type` argument. We show here the use of the `format_output_H` function to extract the `S` compartment: 

```{r, fig.height = 6, fig.width = 8}
S_output <- format_output_H(r1, var_select = "S")
head(S_output)
```

By using the `keep` argument we can extract the value of these variables for each of the 21 patches or each of the two vaccine status (1 = non vaccinated, 2 = vaccinated) of the model. By setting `keep` equal to `patch`, the output has an additional column for the patch:

```{r, fig.height = 6, fig.width = 8}
S_output <- format_output_H(r1, var_select = "S", keep = "patch")
head(S_output)
```

By specifing the same arguments `var_select` and `keep` arguments in the plot function we plot the `S` compartment by patch:

```{r, fig.height = 6, fig.width = 8}
plot(r1, type = "H", var_select = "S", keep = "patch")
```

We can also plot the mosquito compartments of the model using the same `plot` function. To do that we modify the value of the `type` argument.

```{r, fig.height = 6, fig.width = 8}
plot(r1, type = "M")
```

This plots the following mosquito compartments:

* `Mwt_S` - Susceptible wild type adult 
* `Mwt_E1` - Exposed wild type in incubation stage 1
* `Mwt_E2` - Exposed wild type in incubation stage 2
* `Mwt_I1` - Infectious wild type
* `Mwb_S` - Susceptible wolbachia-infected adult 
* `Mwb_E1` - Exposed wolbachia-infected in incubation stage 1
* `Mwb_E2` - Exposed wolbachia-infected in incubation stage 2
* `Mwb_I1` - Infectious wolbachia-infected

As a sanity check, we also plot the variables which are directly modified by seasonal forcing, namely mosquito larve carrying capacity (`Kc`), extrinsic incubation period (`eip`) and adult mosquito mortality rate (`Delta`), and the reproduction number ($R_0$). We can make use of the function `format_output_M()`, which allows to format the model outputs. `format_output_M()` by default extracts the mean value of `Kc`, `eip` and `Delta` across patches. Once we have extracted the model output we can write our own plotting code: 

```{r, fig.height = 6, fig.width = 8}
library(ggplot2)
library(patchwork)

Kc <- format_output_M(r1, var_select = "Kc")
Kc_p <- ggplot(Kc, aes(x = t, y = y)) +
  geom_line(color = 'royalblue', size = 0.5) +
  scale_x_continuous("Time") +
  scale_y_continuous("Mean across patches") +
  ggtitle("Mosquito larvae carrying capacity") +
  theme_bw()

eip <- format_output_M(r1, var_select = "eip")
eip_p <- ggplot(eip, aes(x = t, y = y)) +
  geom_line(color = 'royalblue', size = 0.5) +
  scale_x_continuous("Time") +
  scale_y_continuous("Mean across patches") +
  ggtitle("Extrinsic Incubation Period") +
  theme_bw()

delta <- format_output_M(r1, var_select = "Delta")  
delta_p <- ggplot(delta, aes(x = t, y = y)) +
  geom_line(color = 'royalblue', size = 0.5) +
  scale_x_continuous("Time") +
  scale_y_continuous("Mean across patches") +
  ggtitle("Adult mosquito daily mortality rate") +
  theme_bw()

Rt <- format_output_M(r1, var_select = "R0t_1")  
Rt_p <- ggplot(Rt, aes(x = t, y = y)) +
  geom_line(color = 'royalblue', size = 0.5) +
  scale_x_continuous("Time") +
  scale_y_continuous("Mean across patches") +
  ggtitle("Time varying reprodution number") +
  theme_bw()

all <- Kc_p + eip_p + delta_p + Rt_p
all
```
A expected the mean value of `Kc`, `eip` and `Delta` does no change throughout the simulation when seasonality is turned off.
