---
title: "Run a seasonal model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run a seasonal model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
library(ZikaModel)
```

## Seasonality 
The model allows to account for the effect of seasonal variations in climatic 
variables (e.g. temperature and precipitation) on Zika transmission dynamics. 
In the model, seasonality affects: 

* adult mosquitoes mortality;
* mosquito larvae carrying capacity;
* Extrinsic Incubation Period.

A deterministic model which includes the effect of seasonality can be implemented
with the following:

```{r}
# create a vector of human age groups 
age_init <- c(1, 9, 10, 10, 10, 10, 10, 10, 10, 10, 10)
  
# create a vector of human mortality rates 
deathrt <- c(1e-10, 
             1e-10, 
             1e-10, 
             0.00277068683332695, 
             0.0210680857689784,
             0.026724997685722,
             0.0525354529367476,
             0.0668013582441452,
             0.119271483740379,
             0.279105747097929,
             0.390197266957464)
             
# provide the length of time (in days) that you want to run the model for
time_frame <- 364 * 100

model_run <- ZikaModel::run_model(agec = age_init,
                                  death = deathrt,
                                  nn_links,
                                  time = time_frame,
                                  season = TRUE)
```

The model results can be visually inspected with

```{r, fig.height = 5, fig.width = 5}
model_run$proportions
```

```{r, fig.height = 6, fig.width = 8}
model_run$diagnostics[[1]]
```

```{r, fig.height = 6, fig.width = 8}
model_run$diagnostics[[2]]
```
