---
title: "Run a seasonal model"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Run a seasonal model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, echo=FALSE}
library(ZikaModel)
```

## Seasonality 
The model allows to account for the effect of seasonal variations in climatic 
variables (e.g. temperature and precipitation) on Zika transmission dynamics. 
In the model, seasonality affects: 

* adult mosquitoes mortality;
* mosquito larvae carrying capacity;
* Extrinsic Incubation Period.

A deterministic model which includes the effect of seasonality can be implemented
with the following:

```{r}
seasonal_model_run <- ZikaModel::run_model(agec = age_init,
                                           death = deathrt,
                                           nn_links,
                                           time = time_frame,
                                           season = TRUE)
seasonal_model_run$plot
```
